The CMS Data Aggregation System

### 2010-05-04 doc/paper.tex
Work on Introduction and Related Work sections
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1805 4525493e-7705-40b1-a816-d608a930855b

%\cortext[cor1]{Corresponding autho}

\begin{abstract}
Meta-data plays significant role in large modern enterprises, 
research experiments and digital libraries where it comes from a different 
sources and distributed in a variety of forms and digital formats. 
It is organized and managed by constantly evolving software using 
both relational and non-relational data sources. Even though we can apply
information retrieval approach to non-relation data sources,
we can't do so for relational ones, where information is accessed via
pre-established set of data-services.

Here we discuss a new data aggregation system which consumes, 
indexes and delivers information from different relational and 
non-relational data sources to answer cross data-service queries 
and explore meta-data associated with petabytes of experimental data. 
We combine simplicity of keyword-based search with precision of RDMS
under the new system. The aggregated information being collected from various sources,
allowing end-users to place dynamic queries, get precise answers and 
trigger information retrieval on demand. Such close to real-time system 
based on studies  of use cases of the CMS particle physics experiment at 
the LHC where we used a large scale, distributed computing environment 
to motivate this work.
\end{abstract}

@@

%% main text
\section{Introduction}
The European Organization for Nuclear Research, known as CERN, plays a leading
role in fundamental studies of physics. But it is also known as a place where
many innovations in the area of computer science were developed, e.g. World Wide Web.
Today, the Large Hadron Collider (LHC) at CERN is marking a new era of High Energy
Physics (HEP), promising to deliver a few PB of data each year. 
At this scale the ``information discovery'' within heterogeneous, distributed 
environment becomes a key ingredient of successful data analysis.
The data and associated meta-data are produced in variety of forms and digital formats.
They are stored and retrieved from relational and non-releational data-sources, such as 
RDMS systems, document oriented databases, blogs, twikies, file systems,
customized applications, etc. 
Working in such environment requires a lot of expertise and our users, physicists, 
are always looking for simple, intuitive and flexible
tool(s) to look-up their desired data. A well-known solutions, such as data-services
and search engines are tighted to a specific data sources and end-users are left 
with a manual task to bookkeep and relatate information from them.

Here we present a work on Data Aggregation System (DAS) designed for
CMS High-Energy Experiment at LHC which provides
ability to query, search and aggregate information from different 
data-services and trigger information retrieval on demand.

The rest of this paper is organized as following. 
In section \ref{RelatedWork} we discuss related work in a domain of 
keyword search over relational data sources.
Section \ref{DataModel} describes CMS experiment and its data model. In section
\ref{DAS} we present architecture of the DAS system, including discussion of its
various components. Finally, our results are summarized in section \ref{Results}.

@@

\section{Related Work\label{RelatedWork}}
Even though the idea of keyword queries over the relational database
%querying relation databases via keyword based search
%algorithms 
is not knew it is still under discussion in computer science domain.
A few alternative solutions has been proposed to address this issue
in last several years. In \cite{DBXplorer} the conjunctive keyword queries,
i.e. retrieval of documents that contain all query keywords, has
been discussed. It requires that all keywords are matched in a row from a
single table or joined tables. Authors of \cite{QueryAnswer} made a step forward
by considering information around matched values and presents customized answers to
end-users over the generated new database. These and other related 
works, see \cite{DBXplorer, QueryAnswer} for more references,
are based on some kind of the graph over the underlying database and
use a generated database out of pristine one to perform the work.
The proximity of the results in those techniques 
are mostly confined to a text-based values stored in a database, since 
usage of numerical values in the input search field is mostly useless without 
knowledge of the context of the value. For example, typing a value 100 
in the search input field leads to plenty of inappropriate matches, 
such as row ids, and in order to be
answered correctly requires additional keyword, clarifying its meaning, 
which should be matched with particular
entity (table.column) in the underlying database.
Therefore to answer presicely the logical questions, such as  
{\it find me total number of files whose size more then 20 but less then 
100} would be hard or almost impossible in \cite{DBXplorer, QueryAnswer}, 
while it can be easily accomplished using SQL syntax.

To address those limitations an attempt to build a simple, intuitive and 
flexible query language was introduced in \cite{DBS-QL}.
It represents a power of SQL while hiding underlying relational schema from 
the end-users. As a results a human questions were intuitively mapped into 
simple queries. For instance, the question
{\it I'm looking for files who contain data taken on certain date and located at
particular site} was represented as simple as \cite{DBS-QL}
\begin{verbatim}
find file where date 2009-01-02 11:59 CET and site = T2
\end{verbatim}
This solution represents a common use cases within HEP community and provide
intuitive mapping between mental model of end-users and underlying database schema.

At the same time the data aggregation is commonly addressed in the context of
federated databases, see for example \cite{FedDB}. It unifies data coming 
from different RDMS systems into the federated DB where SQL queries can be used to 
search desired data. Even though it would be interesting to apply technique 
discussed in \cite{DBS-QL} to a federated DB it still requires 
an access to the underlying database schema. This obviously creates a problem
in distributed, heterogeneous environment where different security policies
are in place.

To avoid those limitations we decided to build a keyword-search based system around
existing data-services whose nature and access policies are defined by developers.
Keeping in mind that meta-data can come in different forms and data formats,
we used a document-oriented database back-end with full support of the query language.
The key ingredients were the setup of the mapping and analytics services which provide
necessary glue across existing API/notations and user queries. This approach 
allows us to perform dynamic data aggregation across various data-services, achieve
precise answers using keyword-based queries
and organize information retrieval on demand.
%to store meta-data information 
%We wanted to expand this approach further and apply it 
%to handle cross data-service queries from broad variety 
%of relational and non-relational data-sources within 
%our distributed heterogeneous environment.

\section{CMS data model\label{DataModel}}
The CMS, Compact Muon Solenoid, \cite{CMS} 
is one of the two large general-purpose particle physics detectors built on 
the proton-proton Large Hadron Collider (LHC) at CERN in Switzerland and France. 
It is designed to explore the frontier of the HEP and provide physicists
the ability to look at conditions presented at the early stage of our Universe.
More then 3000 physicists from 183 institutions reprepresenting almost 
38 countries are involved in the design, construction and maintenance of the experiment.

The CMS distributed computing and data model \cite{CMSDataModel} 
is designed to process
and efficiently manage a few PBs of data expected each year
in CMS operation at LHC. The computing resources provided by members of CMS
collaboration are geographically distributed, 
interconnected via high throughput networks and operated by means 
of Grid software. The model allows to cover broad variety of
hardware, mass and storage elements and configuration of the
clusters. To accomodate efficient data processing CMS uses
a multi-Tier distributed model, where specific tasks of data taking,
processing, archival and distribution are assigned to each tier based
on CMS data model. For example, the Tier-0 center at CERN is responsible
for archiving the data coming out from the detector, prompt first pass reconstruction
and data distribution to Tier-1 centers, located around the world.
%The 7 Tier-1 centers
%located in France, Germany, Ialy, Spain, Taiwan, United Kingdom and United States
%keep a portion (copy) of the data delivered by Tier-0 for further processing.
They provide storage and CPU power for high priority analysis jobs.
The Tier-2 centers
%, located at more then 50 sites around the world,
are dedicated for user analysis tasks and production of simulated data.

@@

\section{Data Aggregation System\label{DAS}}
The goal of the Data Aggregation System was to combine and
aggregate meta-data from existing data-services
under one umrella, utilizing their access rules, APIs, security policies, etc., and
allowing users to perform queries against them. In contrast, the approaches
discussed in \cite{DBXplorer, QueryAnswer, FedDB} are
enforced a publication step of data into a dedicated database 
or central repositories \cite{iRODS}. Therefore, quite often it is required
that data should be converted into uniform data formats in order enable
aggregation, see \cite{OpenArchive}. In DAS architecture the data are provided
by existing data-services and infrastructure, whose implementation,
data representation and security policies are left intact. Consequently,
DAS does not require a data preservation and transaction capabilities. 
Instead DAS can be treated
as proxy or caching layer where information retrieval are triggered dynamically
by user queries. Such infrastructure allows data sharing across different queries and
data aggregation using appropriate mapping. Below we discuss in details the
DAS architecture.

\subsection{DAS architecture}
The design of the DAS system was based on previous studies of the CMS Data 
Bookkeeping System (DBS) \cite{DBS, DBS07}. The DBS Data Discovery service
\cite{DD} was based on DBS Query Language (DBS-QL) \cite{DBS-QL}. It provided
a syntax similar to SQL, without specific knowledge of underlying DB schema.
%We carefully analyzed user
%requests and queries placed into the DBS, their patterns, frequencies and latency. 
%As a result we created a data discovery service \cite{DD}. Originally it used MVC
%architecture and later, being re-factored as pure presentation layer on top of 
%the DBS Query Language (DBS-QL) \cite{DBS-QL} with support of dynamic queries against
%DBS back-ends. The DBS-QL \cite{DBS-QL} syntax was based on SQL query language, where 
%we took away relations and details on underlying schema from end-users. 
%This was achieved by using Dijkstras shortest path algorithm 
%to establish necessary table joins based on provided information in user queries and
%knowledge of foreign key relationships from auto-loaded database schema.
%The DBS-QL keys exposed to end-users were choosen based on a jargon used by
%physicists. They're internally mapped into database entitites and attributes, 
%e.g. the {\it run} DBS-QL key refers to run number attribute of run summary table.
A quick adoption and wide usage of DBS-QL in CMS gave us confidence in choosen approach 
and provide a basis for building DAS system.

As was pointed out in \cite{Arms} a mixed content and 
mixed meta-data and meta-data consistency should be considered as a whole in design 
of the system to successful information discovery. 
Starting from this ground we designed DAS system as an
additional layer on top of the existing data-services
within CMS computing infrastructure by imposing the following set of requirements:
\begin{itemize}
\item support a keyword based search queries with ability to use condition operators;
%\item pluggable interface to existing and up-coming data-service APIs;
\item use existing heterogeneous software environment and distributed nature of the data-services;
\item preserve access rules and security policies of individual data-services;
%\item be flexible to adjust to dynamic nature of data-service APIs, 
%due to continuously changing detector and software environments;
\item retrieve data on demand and aggregate them if necessary across
multiple data-services;
\item be transparent to data-services implementation and their data formats;
%\item support legacy data-services, applications and APIs.
\end{itemize}
\begin{figure}[htb]
\centering
\includegraphics[width=100mm]{DAS_Cache_and_Analytics.pdf}
\caption{
The DAS architecture diagram. 
Input query were placed into analytics DB and mapped into a set of
data-service APIs and their parameters. If superset of the query has been
found in DAS cache, the appropriate results were delivered back to the users.
Otherwise a data-service APIs were invoked, triggering DAS cache population. 
Once data become available and aggregated in DAS cache, users were able 
to see the results. This on demand system was supplemented by DAS robots 
(unix daemons) who were consulting DAS analytics database to enable data 
population into DAS cache for most common query requests.
}
\label{DAS_cache}
\end{figure}

\noindent
Based on this set we rule out the choice of RDMS as DAS back-end(s)
for several reasons. We agreed to support dynamic queries, data
aggregation on demand and pluggable data-services. Therefore a choice 
of concrete schema for data storage was problematic. Moreover we didn't impose
any transactions and data persistency in DAS. This allows us to look 
for alternative IT solutions.
Through the analysis of available options, such as a file-based and memory caches, 
key-value databases, documented-oriented databases we made our choice in favor 
of the last technology. Among them we evaluated CouchDB \cite{CouchDB} and 
MongoDB \cite{MongoDB}. Both systems provide schema free document
storage, replication and fail-over features, but our choice in favor of 
MongoDB was obvious, due to its support of dynamic queries, 
full indexies, including inner objects and embedded arrays,
and auto-sharding. Our prelimiary benchmarks shown that it can sustain
the desired load and size of our meta-data information. We used 
document-oriented database, MongoDB, as a back-end for the three 
main DAS components: Mapping, Analytics and DAS cache databases, 
which will be discussed next.

@@

\subsection{DAS queries}
Apart from the previous DBS-QL syntax we decided to drop its syntax and allow users
express their queries in a free text-based form. Although we must clarify its
meaning. We used a pre-defined set of keys registered in DAS to describe common
entities of our data. For example, \verb+dataset+, 
\verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. Their choice
was based on analysis of daily work operations of our users, see \cite{DBS07}.
Each key had a set of attributes with primary one assigned to the key itself.
For intance, the \verb+block+ key had the following attributes
\verb+name+, \verb+size+, \verb+replica+ etc., 
with primary attribute being \verb+name+.
The choice of attributes were based on a set of meta-data information
associated with a given DAS key. Therefore end-users where able to specify
a context in their queries by using appropriate attribute(s). For example,
for query \verb+block+ we didn't perform a string matching to the
``block'' string, but rather look-up a block objects in a DAS, while query
\verb+block.name=abc*+ instructs system to look-up appropriate blocks
with matched names. At the same time, to complement the stregth of the 
DAS, we do plan to index dynamically retrieved meta-data and add a keyword-based 
search later.

@@

%We used RESTful model \cite{REST} for all DAS cache server APIs.
%It allows us to share core library and other components during development of 
%various DAS clients, e.g. CLI, web interface, robots, etc. and hides
%complexity of access to individual data-services and their security policies
%as well as proper organize caching solution.
%For instance, to access on-line Run Summary information we were forced to use
%Grid certificates. Instead of propagating user certificate to data-service
%the DAS cache server works as a proxy to this server and used host certficiate 
%to serve such requests on behalf of the users.

\section{Results\label{Results}}
In this section we discuss preliminary restults achieved with DAS system
using the following CMS data-services:
\begin{itemize}
\item Data Bookkeeping System, DBS \cite{DBS}, the CMS meta-data service;
\item Physics Experiment Data Export, PhEDEx \cite {PhEDEx}, the 
CMS file transfer system;
\item SiteDB \cite{SiteDB}, the CMS site resources data-service;
\item RunSummary \cite{RunSummary}, the CMS run and trigger data-service;
%\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
%\item DQ \cite{DQ}, Data Quality data-service which collects information
%about detector conditions during data taking;
%\item Overview \cite{Overview}, collects information about CMS
%transfer rate, etc.;
\item Dashboard \cite{Dashboard}, the LHC virtual organization
monitoring data-service portal;
\end{itemize}
Each data-service has its own scope, size and peculiar properties. 
For instance, the data in PhEDEx are transient due to constant 
migration of CMS data; the DBS system were divided into dozen 
of individual instances with own size and lifetime, who has been used by different production 
teams for various bookkeeping tasks; the RunSummary data-service
operates on private network accessible at CMS online operation center.
%Table \ref{db_size} shows the current size of meta-data from aforementioned 
%data-services.
%\begin{figure}[htb]
%\centering
%\includegraphics[width=100mm]{db_size.pdf}
%\caption{
%Current size of meta-data for CMS data-services participated in DAS, before
%official data-taking. 
%}
%\label{db_size}
%\end{figure}

@@

%Size of  & Dashboard & DBS & PhEDEx & SiteDB & RunSummary \\
%\hline
%tables (GB) & 40.8  & 28.0 & 5.5 & 0.006 & 0.9 \\
%indexies (GB) & 135.9 & 57.2 & 12.6 & 0.004 & 2.7 \\
%\hline
%\hline
%\end{tabular}
%\caption{
%Current size of meta-data for CMS data-services participated in DAS, before
%official data-taking. 
%}
%\label{db_size}
%\end{table*}
At the moment, the amount of meta-data stored in all services is equal to
$\sim$75 GB and $\sim$200 GB in tables and index sizes, respectively. While 
we estimate to accumulate $\sim$500 GB of the meta-data each year during 
normal CMS operation at LHC.
%\footnote{This number is based on tests 
%performed in CMS during Cosmic Run at Full Tesla (CRAFT) data taking \cite{CRAFT09}.}
%In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
%billion events, 330TB in 32 datasets:
%AlCa/Calibration RAW: 926 million events, 29TB
%Physics RAW: 1.24 billion events, 147TB
%PromptReco: 1.24 billion events, 111TB
%Bulk ALCARECO: 2.34 billion events, 11TB
%Express FEVT: 105 million events, 24TB
%Express ALCARECO: 153 million events, 0.99TB
%HLTMON FEVTHLTALL: 31 million events, 7.4TB
%Above event numbers have signicant overlap by design, 
%2.2 billion unique events, including 524 million cosmic muon
%triggers
% 0.3PB - 50GB meta-data in DBS
% 3PB/y - x GB

@@

\begin{table*}[hbt]
\centering
\begin{tabular}{llllll}\hline
\hline
System & Format & Size & Records & Elapsed time & Elapsed time \\
& & & & no cached data & w/ cached data \\
\hline
DAS (DBS) & XML & 187MB & 368716 & 53 sec & 0.92 sec \\
DAS (PhEDEx) & XML & 164MB & 175884/170131 & 160 sec & 0.94 sec \\
DAS (aggregate) & JSON & 357MB & 374468 & 256 sec & 44.8 sec \\
\hline
\hline
\end{tabular}
\caption{Time required to fetch, parse and aggregate block information
from DBS and PhEDEx systems. In the case of PhEDEx
system we show total number of fetched records 175884 together with
number of records merged in DAS 170131. The total number of DAS records 
were calculated as number of aggregated records plus left overs,
the non-matched records from both systems.}
\label{DAS_benchmark}
\end{table*}

>>>>

%\cortext[cor1]{Corresponding author}

\begin{abstract}
Metadata plays a significant role in large modern enterprises, 
research experiments and digital libraries where it comes from many different 
sources and is distributed in a variety of digital formats. 
It is organized and managed by constantly evolving software using 
both relational and non-relational data sources. Even though we can apply
an information retrieval approach to non-relation data sources,
we can't do so for relational ones, where information is accessed via
a pre-established set of data-services.

Here we discuss a new data aggregation system which consumes, 
indexes and delivers information from different relational and 
non-relational data sources to answer cross data-service queries 
and explore metadata associated with petabytes of experimental data. 
We combine the simplicity of keyword-based search with the precision of RDMS
under the new system. The aggregated information is collected from various sources,
allowing end-users to place dynamic queries, get precise answers and 
trigger information retrieval on demand. Based on the use cases of the CMS experiment, 
we have performed a set of detailed, large scale tests the results of which 
we present in this paper.
\end{abstract}

@@

%% main text
\section{Introduction}
The European Organization for Nuclear Research, known as CERN, plays a leading
role in fundamental studies of physics. It is also known as a place where
many innovations in the area of computer science were developed, e.g. World Wide Web.
Today, the Large Hadron Collider (LHC) at CERN is marking a new era of High Energy
Physics (HEP), promising to deliver a few PB of data each year. 
At this scale the ``information discovery'' within heterogeneous, distributed 
environment becomes a key ingredient of successful data analysis.
The data and associated meta-data are produced in variety of forms and digital formats.
They are stored and retrieved from relational and non-relational data-sources, such as 
RDMS systems, document oriented databases, blogs, twikies, file systems,
customized applications, etc. Working in such environment requires a lot 
of expertise and easy way to look-up desired information is always a necessity.
A well-known solutions, e.g. data-services, are tighten to a specific data 
sources and end-users are left with a manual task to bookkeep and 
relate information from them.

Here we present a work on Data Aggregation System (DAS) which provides
ability to search and aggregate information across different 
data-services. While it was designed for CMS High-Energy Physics 
experiment at LHC, it can be applied elsewhere.

The rest of this paper is organized as following. 
In section \ref{RelatedWork} we discuss related work in a domain of 
keyword search over relational data sources.
Section \ref{DataModel} describes the CMS experiment and its data model. In section
\ref{DAS} we present architecture of the DAS system, including discussion of its
various components. Finally, our results are summarized in section \ref{Results}.

@@

\section{Related Work\label{RelatedWork}}
Even though the idea of keyword queries over the relational database
%querying relation databases via keyword based search
%algorithms 
is not knew it is still under discussion in computer science domain.
A few alternative solutions has been proposed to address this issue
in last several years. In \cite{DBXplorer} the conjunctive keyword queries,
i.e. retrieval of documents that contain all query keywords, has
been discussed. It requires that all keywords are matched in a row from a
single table or joined tables. Authors of \cite{QueryAnswer} are made a step forward
and evaluated information around matched values. The results were presented as
customized templates to end-users. 
These and other related works, see \cite{DBXplorer, QueryAnswer} 
for more references, are based on some kind of the graph over the underlying 
database and generated a new database out of pristine one to perform the work.
The proximity of the results in those techniques are mostly confined to a 
text-based values stored in a database, since the usage of numerical ones 
is mostly useless without knowledge of the context of the input value. 
For example, typing a value 100 in the search input field can lead to 
plenty of inappropriate matches, e.g. row ids. In order to be
answered correctly it requires additional keyword, clarifying its meaning, 
which should be matched with particular entity (table.column) in the 
underlying database. Therefore the logical questions, e.g.
{\it find me total number of files whose size more then 20 but less then 
100} are hard to answer precisely using approaches outlined in \cite{DBXplorer, QueryAnswer}, 
while it can be easily accomplished using SQL.

To address those limitations a keyword-based query language was introduced in \cite{DBS-QL}.
It used a set of pre-defined keys mapped to underlying schema and shortest path
algorithm which employed a foreign key constrains to build SQL queries. As a result
the question
{\it I'm looking for files who contain data taken on certain date and located at
particular site} was represented as simple as \cite{DBS-QL}
\begin{verbatim}
find file where date 2009-01-02 11:59 CET and site = T2
\end{verbatim}
without specifying any relational details.
This solution represents a common use case in HEP community and provides
an intuitive mapping between mental model of the end-users and 
underlying database schema.

At the same time the data aggregation is commonly addressed in the context of
federated databases, see for example \cite{FedDB}. The data from different 
RDMS systems are stored into the federated DB where SQL queries can be used to 
search desired data. Apart from many internal obstacles, e.g.
schema, domain and naming conflicts, the external ones are often
a barrier to work with in distributed, heterogeneous environment where
different security policies are in place. At the end an usage of SQL and/or
any other Query Language will be required to search for data.
%Even though it would be interesting to apply technique 
%discussed in \cite{DBS-QL} to a federated DB it still requires 
%an access to the underlying database schema.

To avoid those limitations we decided to build a keyword-search based system around
existing data-services whose nature and access policies are intact.
%The information retrieval were triggered on demand and stored into the cache.
The key ingredients were the setup of the mapping and analytics services which provided
a necessary glue across existing API/notations and user queries. This approach 
allows us to perform dynamic data aggregation across various data-services, achieve
precise answers using keyword-based queries
and organize information retrieval on demand.
%to store meta-data information 
%We wanted to expand this approach further and apply it 
%to handle cross data-service queries from broad variety 
%of relational and non-relational data-sources within 
%our distributed heterogeneous environment.

\section{CMS data model\label{DataModel}}
The CMS, Compact Muon Solenoid, \cite{CMS} 
is one of the two large general-purpose particle physics detectors built on 
the proton-proton Large Hadron Collider (LHC) at CERN in Switzerland and France. 
It is designed to explore the frontier of the HEP and provide physicists
the ability to look at conditions presented at the early stage of our Universe.
More then 3000 physicists from 183 institutions representing almost 
38 countries are involved in the design, construction and maintenance of the experiment.

The CMS distributed computing and data model \cite{CMSDataModel} 
is designed to process
and efficiently manage a few PBs of data expected each year
in CMS operation at LHC. The computing resources provided by members of CMS
collaboration are geographically distributed, 
interconnected via high throughput networks and operated by means 
of Grid software. The model allows to cover broad variety of
hardware, mass and storage elements and configuration of the
clusters. To accommodate efficient data processing CMS uses
a multi-Tier distributed model, where specific tasks of data taking,
processing, archival and distribution are assigned to each tier based
on CMS data model. For example, the Tier-0 center at CERN is responsible
for archiving the data coming out from the detector, prompt first pass reconstruction
and data distribution to Tier-1 centers, located around the world.
%The 7 Tier-1 centers
%located in France, Germany, Italy, Spain, Taiwan, United Kingdom and United States
%keep a portion (copy) of the data delivered by Tier-0 for further processing.
They provide storage and CPU power for high priority analysis jobs.
The Tier-2 centers
%, located at more then 50 sites around the world,
are dedicated for user analysis tasks and production of simulated data.

@@

\section{Data Aggregation System\label{DAS}}
The goal of the Data Aggregation System was to combine and
aggregate meta-data from existing data-services
under one umbrella, utilizing their access rules, APIs, security policies, etc., and
allowing users to perform queries against them. In contrast, the approaches
discussed in \cite{DBXplorer, QueryAnswer, FedDB} are
enforced a publication step of data into a dedicated database 
or central repositories \cite{iRODS}. Therefore, quite often it is required
that data should be converted into uniform data formats in order enable
aggregation, see \cite{OpenArchive}. In DAS architecture the data are provided
by existing data-services and infrastructure, whose implementation,
data representation and security policies are left intact. Consequently,
DAS does not require a data preservation and transaction capabilities. 
Instead DAS can be treated
as proxy or caching layer where information retrieval are triggered dynamically
by user queries. Such infrastructure allows data sharing across different queries and
data aggregation using appropriate mapping. Below we discuss in details the
DAS architecture.

\subsection{DAS architecture}
The design of the DAS system was based on previous studies of the CMS Data 
Bookkeeping System (DBS) \cite{DBS, DBS07}. The DBS Data Discovery service
\cite{DD} was based on DBS Query Language (DBS-QL) \cite{DBS-QL}. It provided
a syntax similar to SQL, without specific knowledge of underlying DB schema.
%We carefully analyzed user
%requests and queries placed into the DBS, their patterns, frequencies and latency. 
%As a result we created a data discovery service \cite{DD}. Originally it used MVC
%architecture and later, being re-factored as pure presentation layer on top of 
%the DBS Query Language (DBS-QL) \cite{DBS-QL} with support of dynamic queries against
%DBS back-ends. The DBS-QL \cite{DBS-QL} syntax was based on SQL query language, where 
%we took away relations and details on underlying schema from end-users. 
%This was achieved by using Dijkstra's shortest path algorithm 
%to establish necessary table joins based on provided information in user queries and
%knowledge of foreign key relationships from auto-loaded database schema.
%The DBS-QL keys exposed to end-users were chosen based on a jargon used by
%physicists. They're internally mapped into database entities and attributes, 
%e.g. the {\it run} DBS-QL key refers to run number attribute of run summary table.
A quick adoption and wide usage of DBS-QL in CMS gave us confidence in chosen approach 
and provide a basis for building DAS system.

As was pointed out in \cite{Arms} a mixed content and 
mixed meta-data and meta-data consistency should be considered as a whole in design 
of the system to successful information discovery. 
Starting from this ground we designed DAS system as an
additional layer on top of the existing data-services
within CMS computing infrastructure by imposing the following set of requirements:
\begin{itemize}
\item support a keyword based search queries with ability to use condition operators;
%\item pluggable interface to existing and up-coming data-service APIs;
\item use existing heterogeneous software environment and distributed nature of the data-services;
\item preserve access rules and security policies of individual data-services;
%\item be flexible to adjust to dynamic nature of data-service APIs, 
%due to continuously changing detector and software environments;
\item retrieve data on demand and aggregate them if necessary across
multiple data-services;
\item be transparent to data-services implementation and their data formats;
%\item support legacy data-services, applications and APIs.
\end{itemize}
\begin{figure}[htb]
\centering
\includegraphics[width=100mm]{DAS_Cache_and_Analytics.pdf}
\caption{
The DAS architecture diagram. 
Input query were placed into analytics DB and mapped into a set of
data-service APIs and their parameters. If superset of the query has been
found in DAS cache, the appropriate results were delivered back to the users.
Otherwise a data-service APIs were invoked, triggering DAS cache population. 
Once data become available and aggregated in DAS cache, users were able 
to see the results. This on demand system was supplemented by DAS robots 
(Unix daemons) who were consulting DAS analytics database to enable data 
population into DAS cache for most common query requests.
}
\label{DAS_cache}
\end{figure}

\noindent
Based on this set we rule out the choice of RDMS as DAS back-end(s)
for several reasons. We agreed to support dynamic queries, data
aggregation on demand and pluggable data-services. Therefore a choice 
of concrete schema for data storage was problematic. Moreover we didn't impose
any transactions and data persistence in DAS. This allows us to look 
for alternative IT solutions.
Through the analysis of available options, such as a file-based and memory caches, 
key-value databases, documented-oriented databases we made our choice in favor 
of the last technology. Among them we evaluated CouchDB \cite{CouchDB} and 
MongoDB \cite{MongoDB}. Both systems provide schema free document
storage, replication and fail-over features, but our choice in favor of 
MongoDB was obvious, due to its support of dynamic queries, 
full indexes, including inner objects and embedded arrays,
and auto-sharding. Our preliminary benchmarks shown that it can sustain
the desired load and size of our meta-data information. We used 
document-oriented database, MongoDB, as a back-end for the three 
main DAS components: Mapping, Analytics and DAS cache databases, 
which will be discussed next.

@@

\subsection{DAS queries}
Apart from the previous DBS-QL syntax we decided to drop its syntax and allow users
express their queries in a free text-based form. Although we must clarify its
meaning. We used a pre-defined set of keys registered in DAS to describe common
entities of our data. For example, \verb+dataset+, 
\verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. Their choice
was based on analysis of daily work operations of our users, see \cite{DBS07}.
Each key had a set of attributes with primary one assigned to the key itself.
For instance, the \verb+block+ key had the following attributes
\verb+name+, \verb+size+, \verb+replica+ etc., 
with primary attribute being \verb+name+.
The choice of attributes were based on a set of meta-data information
associated with a given DAS key. Therefore end-users where able to specify
a context in their queries by using appropriate attribute(s). For example,
for query \verb+block+ we didn't perform a string matching to the
``block'' string, but rather look-up a block objects in a DAS, while query
\verb+block.name=abc*+ instructs system to look-up appropriate blocks
with matched names. At the same time, to complement the strength of the 
DAS, we do plan to index dynamically retrieved meta-data and add a keyword-based 
search later.

@@

%We used RESTful model \cite{REST} for all DAS cache server APIs.
%It allows us to share core library and other components during development of 
%various DAS clients, e.g. CLI, web interface, robots, etc. and hides
%complexity of access to individual data-services and their security policies
%as well as proper organize caching solution.
%For instance, to access on-line Run Summary information we were forced to use
%Grid certificates. Instead of propagating user certificate to data-service
%the DAS cache server works as a proxy to this server and used host certificate 
%to serve such requests on behalf of the users.

\section{Results\label{Results}}
In this section we discuss preliminary results achieved with DAS system
using the following CMS data-services:
\begin{itemize}
\item Data Bookkeeping System, DBS \cite{DBS}, the CMS meta-data service;
\item Physics Experiment Data Export, PhEDEx \cite {PhEDEx}, the 
CMS file transfer system;
\item SiteDB \cite{SiteDB}, the CMS site resources data-service;
\item RunSummary \cite{RunSummary}, the CMS run and trigger data-service;
%\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
%\item DQ \cite{DQ}, Data Quality data-service which collects information
%about detector conditions during data taking;
%\item Overview \cite{Overview}, collects information about CMS
%transfer rate, etc.;
\item Dashboard \cite{Dashboard}, the LHC virtual organization
monitoring data-service portal;
\end{itemize}
Each data-service has its own scope, size and peculiar properties. 
For instance, the data in PhEDEx are transient due to constant 
migration of CMS data; the DBS system were divided into dozen 
of individual instances with own size and lifetime, who has been used by different production 
teams for various bookkeeping tasks; the RunSummary data-service
operates on private network accessible at CMS online operation center.
%Table \ref{db_size} shows the current size of meta-data from aforementioned 
%data-services.
%\begin{figure}[htb]
%\centering
%\includegraphics[width=100mm]{db_size.pdf}
%\caption{
%Current size of meta-data for CMS data-services participated in DAS, before
%official data-taking. 
%}
%\label{db_size}
%\end{figure}

@@

%Size of  & Dashboard & DBS & PhEDEx & SiteDB & RunSummary \\
%\hline
%tables (GB) & 40.8  & 28.0 & 5.5 & 0.006 & 0.9 \\
%indexes (GB) & 135.9 & 57.2 & 12.6 & 0.004 & 2.7 \\
%\hline
%\hline
%\end{tabular}
%\caption{
%Current size of meta-data for CMS data-services participated in DAS, before
%official data-taking. 
%}
%\label{db_size}
%\end{table*}
At the moment, the amount of meta-data stored in all services is equal to
$\sim$75 GB and $\sim$200 GB in tables and index sizes, respectively. While 
we estimate to accumulate $\sim$500 GB of the meta-data each year during 
normal CMS operation at LHC.
%\footnote{This number is based on tests 
%performed in CMS during Cosmic Run at Full Tesla (CRAFT) data taking \cite{CRAFT09}.}
%In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
%billion events, 330TB in 32 datasets:
%AlCa/Calibration RAW: 926 million events, 29TB
%Physics RAW: 1.24 billion events, 147TB
%PromptReco: 1.24 billion events, 111TB
%Bulk ALCARECO: 2.34 billion events, 11TB
%Express FEVT: 105 million events, 24TB
%Express ALCARECO: 153 million events, 0.99TB
%HLTMON FEVTHLTALL: 31 million events, 7.4TB
%Above event numbers have significant overlap by design, 
%2.2 billion unique events, including 524 million cosmic muon
%triggers
% 0.3PB - 50GB meta-data in DBS
% 3PB/y - x GB

@@

\begin{table*}[hbt]
\centering
\begin{tabular}{lllll}\hline
\hline
System & Format & Records & Elapsed time & Elapsed time \\
& & & no cached data & w/ cached data \\
\hline
DAS (DBS) & XML & 382396 & 63 sec & 0.98 sec \\
DAS (PhEDEx) & XML & 186145/180439 & 164 sec & 0.98 sec \\
DAS (aggregate) & JSON & 388108 & 234 sec & 2.05/50.7 sec \\
\hline
\hline
\end{tabular}
\caption{Time required to fetch, parse and aggregate block information
from DBS and PhEDEx systems. In the case of PhEDEx
system we show total number of fetched records 186145 together with
number of records merged in DAS 180439. The total number of DAS records 
were calculated as number of aggregated records plus left overs,
the non-matched records from both systems. The look-up time is presented in
last column, where for DAS it shows as query look-up time (2.05 sec) 
and time to get all records to disk (50.7 sec).}
\label{DAS_benchmark}
\end{table*}

The initial DAS implementation was done in Python language \cite{Python}
and all tests were performed on x64 Linux platform where
1 CPU was used by DAS server and 1 CPU by MongoDB database server. 
%The CPU utilization
%of DAS server was 67\% in average and RAM usage was below 35MB. 
All tests were performed multiple times to avoid fluctuations. 
The elapsed time measured during those tests consists of the 
following components:
\begin{itemize}
\item[]
{\it
Elapsed time = retrieval time + parsing time + re-mapping time 
        + cache insertion/indexing time 
        + (aggregation time) + (output creation time)
}
\end{itemize}
Here the {\it retrieval time} is a time required to access data from remote data-service,
{\it parsing time} is a time required to read and parse received data, {\it re-mapping time}
is a time required to convert data-service meta-data to DAS records for every object
we parse, {\it cache insertion} and {\it indexing time} represents time spent on DAS 
cache back-end, {\it aggregation time} is required to merge objects into DAS records based
on their common key (\verb+block.name+ in this case) and {\it output creation time}
is a time required to write DAS records to disk. Please note that last two
components, {\it aggregation} and {\it output creation time}, were only applied to
one of the sub-system, either DBS or PhEDEx, depending on insertion order.
% since the first one was injected into
%the fresh database, while the second required the merging step.
